---
title: "Exercises2"
author: "Yongqi WANG, Hangjia ZHAO"
output:
  pdf_document: default
header-includes:
 - \usepackage{amsmath}
#  - \usepackage{blkarray} # use blkarray in latex env
#  - \usepackage{bm} # use blkarray in latex env
#  - \usepackage{physics} # use blkarray in latex env
---

```{r setup, include=FALSE, warning=F}
#knitr::opts_chunk$set(echo = TRUE)
#source("draw_matrix.R")
```

\newcommand{\m}[1]{\mathbf{#1}}
\newcommand{\myoe}[2]{\frac{\mathrm{d} #1}{\mathrm{d} #2}}

## Problem 1

### a)

Since there $20$ unique amino acids, there exist $20^{20}$ unique amino acid sequences for a sequence of $L=20$.

### b) 

Codons consist of $3$ DNA sequences is responsible for a single amino acid sequence, the DNA sequence encoding the amino acids sequence is of length of $20 \times 3 = 60$ and has $60^{4}$ uniques sequences.

## Problem 2

### a)

Let $x, y$ be two binary sequences of length $L$, $P(z_i = 1) = P(z_i = 1) = \frac{1}{2}, z \in \{x, y\},i \in \{0,1,\dots, L\}$.

Let $a, b$ be two DNA sequences of length $L$, $P(c_i = A) = P(c_i = T) = P(c_i = C) = P(c_i = G) = \frac{1}{4}, c \in \{a, b\},i \in \{0,1,\dots, L\}$

**Binary**: $$E(d_H(x, y)) = E(\sum_{i=1}^L 1_{x_i \not = y_i}) = \sum_{i=1}^L E(1_{x_i \not= y_i}) = L \cdot P(x_i \not= y_i) = \frac{L}{2}$$

**Random DNA**: $$E(d_H(a, b)) = E(\sum_{i=1}^L 1_{a_i \not = b_i}) = \sum_{i=1}^L E(1_{x_i \not = y_i}) = L \cdot P(a_i \not= b_i) = \frac{3L}{4} $$

### b)

**Binary**: $\binom{L}{K}$ of sequences at a Hamming distance $K$ since the size of $\cal{A}$ is $2$,  $\binom{L}{2}$ sequences at a Hamming distance of $2$.

**Random DNA**: $\binom{L}{K}\cdot 3^K$ because of the increase size in the alphabet $\cal{A}$ to $4$. 

## Problem 3

### a)

$$
W = \begin{bmatrix}
    f_0q_{00} & f_1q_{10} \\
    f_0q_{01} & f_1q_{11} \\
    \end{bmatrix}
 = \begin{bmatrix}
    f_0q & 1-q \\
    f_0(1-q) & q \\
    \end{bmatrix}
$$
The eigenvalues of the above matrix by solving $det(W-\lambda I) = 0$:

$$
\lambda_{1,2} = \frac{f_0(q+1) \pm \sqrt{(f_0(q+1))^2 - 4f_0(2q-1)} }{2}
$$

### b)

Perron-Frobenius theorem states that 

### c)

Assumed $f_0 = 1.5 > 1, q=0.9$
```{r, message=FALSE, warning=FALSE, results=F, fig.height=3, fig.width=5}
library(matlib)
f0 = 1.5
q = 0.9
parms <- c(f0 = f0, f1 = 1, q_00 = q, q_01 = 1-q, q_10 = 1-q, q_11 = q)
times <- c(0:1000)/25
initconds <- c(a = 0.3, b = 0.7)
M = matrix(c(parms["f0"] * parms["q_00"], parms["f0"] * parms["q_01"], 
             parms["f1"] * parms["q_10"], parms["f1"] * parms["q_11"]),
           2, 2)
eig = eigen(M)
#M
#eig

x_star = eig$vectors[, 1]/sum(eig$vectors[, 1])
x_star

sum(x_star * c(parms["f0"], parms["f1"]))

ivp = solve(eig$vectors, initconds)
Xa = ivp[1] * eig$vectors[1, 1] * exp(eig$values[1] * times) +
ivp[2] * eig$vectors[1, 2] * exp(eig$values[2] * times)
Xb = ivp[1] * eig$vectors[2, 1] * exp(eig$values[1] * times) +
ivp[2] * eig$vectors[2, 2] * exp(eig$values[2] * times)

par(mar = c(4, 4, 2, 0.5)) # margin size
par(mgp = c(2.5, 1, 0)) # axis localtion
par(cex.lab = 1.25) # size of y axis label 
par(mfrow=c(1,2)) # 1x2 fig
plot(times, Xa, xlab = "time", ylab = expression(X), main = "",
     col = "dodgerblue", type = "l", lwd = 2)
lines(times, Xb, col = "gold2", lwd = 2)
legend("topleft", legend=c("0", "1"),
       col=c("dodgerblue", "gold2"), lty=1, lwd = 2)
plot(times, Xa/(Xa + Xb), ylim = c(0, 1), xlab = "time", ylab = expression(X),
     main = "", col = "dodgerblue", type = "l", lwd = 2)
lines(times, Xb/(Xa + Xb), col = "gold2", lwd = 2)
legend("topleft", legend=c("0", "1"),
       col=c("dodgerblue", "gold2"), lty=1, lwd = 2)
```


### d)

By b) we know that the equilibrium point for $f_0 = f_1 = 1$ would be 

$$
\lambda = \frac{(q+1) + \sqrt{(q+1)^2 - 4(2q-1)} }{2}
$$

### e)

$$
\lambda = \frac{2f_0 + \sqrt{4f_0^2 - 4f_0} }{2}
$$

Code is available on github repo:
(https://github.com/wyq977/evolutionary-dynamics-2019)
