---
title: "Exercises1"
author: "Yongqi WANG, Hangjia ZHAO"
output:
  pdf_document: default
#header-includes:
#  - \usepackage{blkarray} # use blkarray in latex env
#  - \usepackage{bm} # use blkarray in latex env
#  - \usepackage{physics} # use blkarray in latex env
---

```{r setup, include=FALSE, warning=F}
#knitr::opts_chunk$set(echo = TRUE)
#source("draw_matrix.R")
```

\newcommand{\m}[1]{\mathbf{#1}}
\newcommand{\myoe}[2]{\frac{\mathrm{d} #1}{\mathrm{d} #2}}

## Problem 1

### a)

The model reaches equilibrium points as the $x_{t+1} = f(x_t) = x_t$, let $x^*$ be the equilibrium points of the system.

Solving $f(x) = rx(1-x) = x$, we have: $x^* = 0$ or $x^* = 1 - \frac{1}{r}$ when $r \not= 0$.

### b) 

The stability of the points can be checked by the gradient of $|f'(x^*)| = |r - 2rx^*|$.

$r = 0.5$, $x^*_1 = 0, |f'(x_1^*)| = 0.5 < 1$ and hence \textbf{attractive and stable}.

$x_2^* = -1$, discarded since $x$ can only take value larger or equal to zero.


$r = 1.5$, $x^*_1 = 0, |f'(x_1^*)| = 1.5 > 1$ and hence \textbf{repelling and unstable}.

$x_2^* = \frac{1}{3}, |f'(x_2^*)| = 0.5 < 1$ and hence \textbf{attractive and stable}.


$r = 2.5$, $x^*_1 = 0, |f'(x_1^*)| = 2.5 > 1$ and hence \textbf{repelling and unstable}.

$x_2^* = \frac{3}{5}, |f'(x_2^*)| = 0.5 < 1$ and hence \textbf{attractive and stable}.


### c)

```{r, message=FALSE, warning=FALSE, results=F, fig.height=2.5, fig.width=7.5}
getRep <- function(init, steps, r) {
  x <- rep(0, steps)
  x[1] <- init
  for (i in 2:steps) {
    x[i] <- r * x[i - 1] * (1 - x[i - 1])
  }
  return(x)
}

steps = 100
x1 = getRep(0.5, steps, 0.5)
x2 = getRep(0.5, steps, 1.5)
x3 = getRep(0.5, steps, 2.5)

par(mar = c(4, 4, 2, 0.5)) # margin size
par(mgp = c(2.5, 1, 0)) # axis localtion
par(cex.lab = 1.25) # size of y axis label 
par(mfrow=c(1,3)) # 1x3 fig


plot(x1, xlab = "step", ylab = "x", main = "r=0.5", col = "dodgerblue",
     pch = 20, ylim = c(0, 1))
abline(h = 0, col = "#ff8c00", lwd = 2)

plot(x2, xlab = "step", ylab = "x", main = "r=1.5", col = "dodgerblue",
     pch = 20, ylim = c(0, 1))
abline(h = 1/3, col = "#ff8c00", lwd = 2)

plot(x3, xlab = "step", ylab = "x", main = "r=2.5", col = "dodgerblue",
     pch = 20, ylim = c(0, 1))
abline(h = 3/5, col = "#ff8c00", lwd = 2)
```


### d) 
### e)

As seen in the figures below, 

```{r, message=FALSE, warning=FALSE, results=F, fig.height=2.5, fig.width=5}
x = getRep(0.5, 100, 3.5)
xstm1 <- x[-length(x)]
xst <- x[-1]
y = getRep(0.5, 100, 3.9)
ystm1 <- y[-length(y)]
yst <- y[-1]

par(mar = c(4, 4, 2, 0.5)) # margin size
par(mgp = c(2.5, 1, 0)) # axis localtion
par(cex.lab = 1.25) # size of y axis label 
par(mfrow=c(1,2)) # 1x2

plot(xstm1, xst, xlab = expression(x[t - 1]), ylab = expression(x[t]),
     main = "r=3.5", col = "dodgerblue", pch = 20, xlim = c(0, 1), 
     ylim = c(0,1))
lines(xstm1, xst, col = "dodgerblue", lwd = 0.5)
abline(b = 1, a = 0, col = "#ff8c00", lwd = 2)

plot(ystm1, yst, xlab = expression(x[t - 1]), ylab = expression(x[t]),
     main = "r=3.9", col = "dodgerblue", pch = 20, xlim = c(0, 1), 
     ylim = c(0,1))
lines(ystm1, yst, col = "dodgerblue", lwd = 0.5)
abline(b = 1, a = 0, col = "#ff8c00", lwd = 2)
```

## Problem 2

### a)

\begin{align*}
\int \frac{d x(t)}{x(t)(1 - \frac{x(t)}{K})} &= \int \lambda dt \quad \text{separation of variables}\\
\int \frac{dx(t)}{x(t)} + \int \frac{dx(t)}{K-x(t)} &= \int \lambda dt \\
\ln{|x(t)|} - \ln{|K - x(t)|} &= \lambda t + C \\
\ln{|\frac{K-x(t)}{x(t)}|} &= -(\lambda t + C) \\
|\frac{K-x(t)}{x(t)}| &= e^{-(\lambda t + C)} \\
\frac{K-x(t)}{x(t)} &= e^{-\lambda t} C_0
\end{align*}

From that we can get:

$$
x(t) = \frac{K}{1+C_0e^{-\lambda t}}, \quad C_0 = \frac{K - x_0}{x_0}
$$
And hence the $x(t) = \frac{Kx_0e^{\lambda t}}{K + x_0(e^{\lambda t} - 1)}$.

### b)

The condition for the equilibrium is that $\myoe{x}{t} = 0$

Solving $f'(x) = 0$, we have: $x^* = 0$ or $x^* = K$.

When $x^* = 0, f'(x^*) = \lambda$ and hence the point is \textbf{stable} if $\lambda <0$ and \textbf{unstable} otherwise.

When $x^* = K, f'(x^*) = -\lambda$ and hence the point is \textbf{unstable} if $\lambda <0$ and \textbf{stable} otherwise.

### c)

```{r, message=FALSE, warning=FALSE, results=F, results=FALSE, fig.height=2.5, fig.width=5}
library(deSolve)
k = 1
lambda = -10
parms <- c()
my.atol <- c(1e-06)
times <- c(0:100)/25
sdiffeqns <- function(t, s, parms) {
  sd1 <- lambda * s[1] * (1 - s[1]/k)
  list(c(sd1))
}

# just below 0
out1m <- lsoda(c(0 - 1e-06), times, sdiffeqns, rtol = 1e-10, atol = my.atol)
# just above 0
out1p <- lsoda(c(0 + 1e-06), times, sdiffeqns, rtol = 1e-10, atol = my.atol)
# just below k
out2m <- lsoda(c(k - 1e-06), times, sdiffeqns, rtol = 1e-10, atol = my.atol)
# just above k
out2p <- lsoda(c(k + 1e-06), times, sdiffeqns, rtol = 1e-10, atol = my.atol)

plot(out1m, xlab = "time", ylab = "x", col = "dodgerblue", lty = 1, lwd = 2, 
     ylim = c(-2, 4), xlim = c(0, 4),
     main = expression(" "*lambda*" = -10, "*K*" = 1"))
lines(out1m, col = "#ff8c00", lty = 3, lwd = 3)
lines(out2m, col = "dodgerblue", lty = 1, lwd = 2)
lines(out2p, col = "#ff8c00", lty = 3, lwd = 3)
```

```{r, message=FALSE, warning=FALSE, results=F, results=FALSE, fig.height=2.5, fig.width=5, echo=F}
library(deSolve)
k = 1
lambda = 10
parms <- c()
my.atol <- c(1e-06)
times <- c(0:100)/25
sdiffeqns <- function(t, s, parms) {
  sd1 <- lambda * s[1] * (1 - s[1]/k)
  list(c(sd1))
}

# just below 0
out1m <- lsoda(c(0 - 1e-06), times, sdiffeqns, rtol = 1e-10, atol = my.atol)
# just above 0
out1p <- lsoda(c(0 + 1e-06), times, sdiffeqns, rtol = 1e-10, atol = my.atol)
# just below k
out2m <- lsoda(c(k - 1e-06), times, sdiffeqns, rtol = 1e-10, atol = my.atol)
# just above k
out2p <- lsoda(c(k + 1e-06), times, sdiffeqns, rtol = 1e-10, atol = my.atol)

out1m <- lsoda(c(0 - 1e-06), times, sdiffeqns, rtol = 1e-10, atol = my.atol)
out1p <- lsoda(c(0 + 1e-06), times, sdiffeqns, rtol = 1e-10, atol = my.atol)
out2m <- lsoda(c(k - 1e-06), times, sdiffeqns, rtol = 1e-10, atol = my.atol)
out2p <- lsoda(c(k + 1e-06), times, sdiffeqns, rtol = 1e-10, atol = my.atol)
plot(out1m, xlab = "time", ylab = "x", col = "dodgerblue", lty = 1, lwd = 2, 
     ylim = c(-2, 4), xlim = c(0, 4),
     main = expression(" "*lambda*" = 10, "*K*" = 1"))
lines(out1m, col = "#ff8c00", lty = 3, lwd = 3)
lines(out2m, col = "dodgerblue", lty = 1, lwd = 2)
lines(out2p, col = "#ff8c00", lty = 3, lwd = 3)
```

## Note

Codes for the solution above is inspired by the tutorials materials and available on [github repo](https://github.com/wyq977/evolutionary-dynamics-2019)